---
title: "Assignment_4"
author: "Nikhila Reddy"
date: "2024-03-16"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(caret)
```

```{r}
#install.packages("factoextra")
library(factoextra)
```

```{r}
library(dplyr)
```

```{r}
library(ggplot2)
library(tidyverse)
```

```{r}
#install.packages("cowplot")
library(cowplot)
```

```{r}
library(readr)
#install.packages("flexclust")
library(flexclust)
```

```{r}
#install.packages("cluster")
library(cluster)
```

```{r}
#install.packages("NbClust")
library(NbClust)
```


```{r}
Pharmaceuticals <- read.csv("C:\\Users\\Nikhi\\Downloads\\Pharmaceuticals.csv")
###to read the given dataset
View(Pharmaceuticals)
###to view the given dataset.
head(Pharmaceuticals)
```
```{r}
###To retrieve the initial observations from the provided dataset.
str(Pharmaceuticals)
```

```{r}
###To examine the format of the provided dataset.
summary(Pharmaceuticals)
```

```{r}
###To examine an overview of the provided dataset.
dim(Pharmaceuticals)
```

```{r}
###To determine the number of rows and columns present in the provided dataset.
colMeans(is.na(Pharmaceuticals))
```

```{r}
row.names(Pharmaceuticals) <- Pharmaceuticals[,2]
Pharmaceuticals <- Pharmaceuticals[,-2]
```


FIRST QUESTION:

a.Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in
conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s)
used, the number of clusters formed, and so on.

```{r}
Pharmaceuticals1 <- Pharmaceuticals[,-c(1,11:13)]
###with exception of "Symbol" and the last three non-numerical variables.
```


NORMALIZING AND CLUSTERING THE DATA

   Here, I have calculated the difference between each observation, and it's important to modify the data beforehand because the default Euclidean distance measure is sensitive to scale.
   
```{r}
norm.Pharmaceuticals1 <- scale(Pharmaceuticals1)
###the data is normalized.
distance <- get_dist(norm.Pharmaceuticals1)
fviz_dist(distance)
```
```{r}
###to measure and plot distance for the given dataset.
```

    The graph illustrates the change in color intensity as distance increases. As expected, the diagonal line     shows a value of zero, indicating the distance between two observations.


#To find the Optimal K value
     
    The Elbow chart and the Silhouette Method are the best techniques for deciding the number of clusters in a k-means model when there are no external factors. The Elbow chart shows that increasing the number of clusters leads to less variation within each cluster, while the Silhouette Method measures how connected an object's cluster is to others.

```{r}
WSS <- fviz_nbclust(norm.Pharmaceuticals1, kmeans, method = "wss")
Silhouette <- fviz_nbclust(norm.Pharmaceuticals1, kmeans, method = "silhouette")
plot_grid(WSS, Silhouette)
```

```{r}
###we used elbow chart and silhouette methods.
set.seed(123)
Kmeans.Pharmaceuticals.Optimalno <- kmeans(norm.Pharmaceuticals1, centers = 5, nstart = 50)
Kmeans.Pharmaceuticals.Optimalno$centers
```

```{r}
Kmeans.Pharmaceuticals.Optimalno$size
```

```{r}
Kmeans.Pharmaceuticals.Optimalno$withinss
```

```{r}
fviz_cluster(Kmeans.Pharmaceuticals.Optimalno, data = norm.Pharmaceuticals1)
```

> Based on the data, we can classify five clusters based on their proximity to the centroids. Cluster 4 exhibits a high Market Capital, while Cluster 2 shows a high Beta, and Cluster 5 has a low Asset Turnover. We can also determine the size of each cluster, with Cluster 1 having the highest number of enterprises, while Cluster 3 contains only two. The sum of squared distances within each cluster provides insights into data dispersion: Cluster 1 (21.9) shows less uniformity compared to Cluster 3 (2.8). Through visualizing the output of the algorithm, we can observe the division of the data into these five distinct groups.

SECOND QUESTION:

b.Interpret the clusters with respect to the numerical variables used in forming the clusters

```{r}
###Employing k-means with k=3 to create clusters.
set.seed(123)
Kmeans.Pharmaceuticals <- kmeans(norm.Pharmaceuticals1, centers =3, nstart = 50)
Kmeans.Pharmaceuticals$centers
```

```{r}
Kmeans.Pharmaceuticals$size
```

```{r}
Kmeans.Pharmaceuticals$withinss
```

```{r}
fviz_cluster(Kmeans.Pharmaceuticals, data = norm.Pharmaceuticals1)
```

> This simplifies the process of identifying and handling the clusters within the analysis. Currently, there are 4 data points allocated to cluster 1, 11 data points assigned to cluster 2, and 6 data points designated to cluster 3.

```{r}
clusplot(norm.Pharmaceuticals1, Kmeans.Pharmaceuticals$cluster, color = TRUE, shade = TRUE, labels = 2,,lines= TRUE)
```

Based on the second graphic, firms in cluster 1 exhibit a low Net Profit Margin and a high Price/Earnings ratio, while those in cluster 2 demonstrate low Asset Turnover and Return on Asset (ROA), yet high Leverage and Estimated Revenue Growth. Cluster 3 did not exhibit any notable characteristics across the parameters we examined.

THIRD QUESTION

c. Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in
forming the clusters)

      By examining the three final categorical variables, namely Median_Recommendation, Location, and Stock Exchange, I aim to identify any patterns in the dataset. I prefer to utilize bar charts to visually represent how firms are distributed across clusters.

```{r}
###dataset is partitioned for last 3 variables.
Pharmaceuticals3 <- Pharmaceuticals %>% select(c(11,12,13)) %>%
  mutate(Cluster = Kmeans.Pharmaceuticals$cluster)
```

```{r}
Median_Rec <- ggplot(Pharmaceuticals3, mapping = aes(factor(Cluster), fill=Median_Recommendation)) +
geom_bar(position = 'dodge') +
labs(x='Clusters', y='Frequency')
Location <- ggplot(Pharmaceuticals3, mapping = aes(factor(Cluster), fill=Location)) +
geom_bar(position = 'dodge') +
labs(x='Clusters', y='Frequency')
Exchange <- ggplot(Pharmaceuticals3, mapping = aes(factor(Cluster), fill=Exchange)) +
geom_bar(position = 'dodge') +
labs(x='Clusters', y='Frequency')
plot_grid(Median_Rec,Location,Exchange)
```


> The graph clearly shows that most of the companies in cluster 3 are situated in the United States and all of them have a 'hold' recommendation for their shares. They are all listed on the New York Stock Exchange. In cluster 2, we opt for 'Moderate Buy' shares, which include only two companies with stocks listed on other exchanges or indexes (AMEX and NASDAQ). Cluster 1 reveals that the four firms are situated in four different countries, and their stocks are traded on the NYSE.

FOURTH QUESTION:

d.Provide an appropriate name for each cluster using any or all of the variables in the dataset.
 
 Here, we can gather all the provided data from the dataset and identify the three distinct groups among the    21 pharmaceutical companies.

 Cluster 1 is termed as 'overvalued international firms' due to the subsequent factors: international           presence, trading on NYSE, low Net Profit Margin, and high Price/Earnings ratio. These companies operate       across multiple continents while raising funds on the world's largest stock exchange (NYSE). They possess      high market valuations that surpass their current earnings levels. To prevent a potential collapse in their    stock prices, they must invest and enhance their earnings to meet investor expectations.

 Cluster 2 is classified as a 'growing and leveraged firm' because of the following attributes: 'Moderate       buy' ratings, low asset turnover and ROA, high leverage, and anticipated revenue growth. Despite their         current low profitability and significant debt, they seem to be highly valued by investors who are willing     to await future growth.

 Cluster 3 is labeled as a 'mature US firm' as it is based in the United States, listed on the NYSE, and        holds 'Hold' ratings.





